---
title: "A synthetic spatial SDM demonstration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A synthetic spatial SDM demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette executes a seeded, synthetic example. It is not a GBIF dataset or an assessment of actual Moroccan pollination services.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(pollinatorSDM)
```

## Fit, select and evaluate separately

```{r spatial-demo}
source(system.file("examples", "spatial_demo.R", package = "pollinatorSDM"))
demo <- run_spatial_demo(seed = 42)
stopifnot(length(intersect(demo$splits$train, demo$splits$validation)) == 0)
stopifnot(length(intersect(demo$splits$train, demo$splits$test)) == 0)
stopifnot(length(intersect(demo$splits$validation, demo$splits$test)) == 0)
demo$metrics
```

Longitude blocks define west training, central validation and east test cells. Predictors and binary labels are generated mathematically; the model learns this artificial relationship. The validation threshold is chosen once and frozen before the final block is evaluated. Preprocessing is not fitted on test rows.

```{r roc, fig.width=6, fig.height=4}
demo$evaluation$roc_plot
```

## Interpretation

The metrics above are generated during this build and describe this synthetic example only. They do not validate species distribution, background-as-absence assumptions or pollination deficit. Real studies need source licences, spatial provenance, source-failure and unit audits, training-only filtering, buffered spatial folds and independent ecological observations.
